\documentclass{article}

\begin{document}
Much has been made of the importance of notions of modularity,
compositionality, and abstraction in representations of knowledge and
computation, especially as it relates to human cognition. 

Modularity is a particular type of relationship between the function
of a system and its structure. A modular system is one in which
various functional requirement of a system are structurally
separated. A paradigmatic example of a modular system is the human
body. The body has various subsystems that perform individual
important tasks: there is a cardiovascular system, an immune system, a
liver, a kidney, a brain, etc. Each one of these systems performs an
individual functions. Small changes to the immune system results in
changes in immune function and small changes in the cardiovascular
system results in changes in cardiovascular structure. And, of course,
it is not the case that changes to one system can't change other
functional aspects of the system; but it seems necessary that the
change be relatively small compared to the change in the associated
functional sytem. In this sense, modularity seems to be analagous to a
linear transformation. Take the structural description of a system to
be a point in some high-dimensional space, where each dimension
corresponds to the value of some structural component. And let the
functional description of a system be another point in a different
high-dimensional space. Then there exists some function between any
two corresponding points. The claim that the system is modular is akin
to saying that the transformation between the structural and
functional descriptions is locally a linear transformation.

The fact that whether a system is modular depends on the mapping
between its structural description and its functional description
means that a system is modular with respect to the granularity of the
functional description. For example, if a car is system that produces
speed given some inputs, then on inspecting the system, we have no
idea whether it is modular or not. 

\end{document}
